---
title: "R Tips"
author: "Kevin Putschko"
date: "September 5, 2018"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
library(tidyverse)
library(rlang)
```

### tibble::deframe()

We talked about **tibble::enframe()**, now we'll talk about its complement, **tibble::deframe()**. The help page tells us the function "converts two-column data frames to a named vector or list, using the first column as name and the second column as value."  

But what is this function actually doing?  It converts a two-column table to a list, assuming the first column contains the names and the second column contains the values of each list element.

For example, let's define a list-table.

```{r}
list_table <- 
  tibble(names  = c("dataframe", "strings", "numbers"),
         values = c(list(mtcars[1:5, ]), list(letters), list(exp(1:10)))
  )

list_table 
```

We see this is a convient way of storing and viewing any number of lists.  But to get this into an actual list form, we use **tibble::deframe()**.

```{r}
list_table %>% deframe()
```



### Use Case

Let's see how this works in practice.  For example, if we have a data dictionary table with at least two columns: Full Name and Short Name.

```{r}

dictionary <- 
  tibble(full  = c("MilePerGallon", "Cylinders", "Weight"),
         short = c("mpg", "cyl", "wt"))

dictionary
```

If we are going to quickly change all the short names to the full names, we can use **tibble::deframe()**.  Using **tibble::deframe()** on dictionary, we see the result is a named vector.

```{r}
dict_deframe <- dictionary %>% deframe()

dict_deframe
```

This next part is a bit advanced, but we are going to use *tidyverse* non-standard evaluation (NSE) to change the names of the columns in the *mtcars* dataframe.

```{r}
mtcars[1:5] %>% glimpse()

# dplyr::rename_() - This method is depreciated in favour of the rlang package methods
mtcars[1:5] %>% rename_(.dots = dict_deframe) %>% glimpse()


# dplyr::rename() with !!! and rlang::syms()
mtcars[1:5] %>% rename(!!! syms(dict_deframe)) %>% glimpse()

```

As you can see, the full column names specified in the data dictionary table were renamed in the table based on the corresponding short names.
