---
title: "R Tips"
author: "Kevin Putschko"
date: "October 1, 2018"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
library(tidyverse)
```

### dplyr::summarise_*

When we want to create conditional summaries, we can use the **dplyr::summarise_*** functions.  These include:

- *summarise_all*: Applies the summary function on all columns
- *summarise_at*: Applies the summary function on specified columns
- *summarise_if*: Applies the summary function if the columns meets a given condition

Let's see how these work in action with the *mtcars* dataframe.

```{r echo=FALSE}
ex_mtcars <- 
  mtcars %>% 
  rownames_to_column() %>% 
  glimpse()

ex_mtcars
```

We see that we have both categorical and numeric columns.  For example, we could count the number of distinct values for all columns.  Traditionally, we could do this one column at a time. As you can see, this can easily get tedious with one row of code for each column.

```{r}
ex_mtcars %>% 
  summarise(n_mpg = n_distinct(mpg),
            n_cyl = n_distinct(cyl),
            n_disp = n_distinct(disp),
            n_hp = n_distinct(hp))
```

The function **dplyr::summarise_all** makes turns this into a single line of code.

```{r}
ex_mtcars %>% summarise_all(n_distinct)
```

If we want different functions for different columns, we can use **dplyr::summarise_at**.

```{r}
cols_categorical <- c("rowname", "cyl", "vs", "am", "gear", "carb")
cols_numeric <- c("mpg", "disp", "hp")

ex_mtcars %>% summarise_at(cols_categorical, n_distinct)
ex_mtcars %>% summarise_at(cols_numeric, mean)

```

This also allows for a list of summary functions to applied to specified columns.

```{r}
funs_numeric <- c("mean", "sd", "length")

ex_mtcars %>% 
  group_by(cyl) %>% 
  summarise_at("mpg", funs_numeric)
```

And in the cases when we are aware of column names in advance, we can use **dplyr::summarise_if**.

```{r}
ex_mtcars %>% summarise_if(is.character, n_distinct)
ex_mtcars %>% summarise_if(is.numeric, mean)
```


